\section{Automated post-training as continual domain adaptation}

The human-learning analogy suggests a closed loop: study, test, diagnose what remains
unknown, build targeted practice, and test again. An unfamiliar domain is not empty;
textbooks, literature, manuals, worked examples, and historical records supply accumulated
knowledge and reference evidence. The model need not know the domain parametrically: it
can read these sources in context and construct curricula with checks from worked answers,
cross-source evidence, or executable procedures. Once verified, continual post-training
internalizes that contextual competence while retaining earlier capabilities. In-context
learning therefore bootstraps the curriculum, while post-training provides continual
learning and domain adaptation rather than another isolated fine-tune.

Figure~\ref{fig:auto-post-training-loop} captures the proposed control loop. Its most
important decision is the reflection step: a failed attempt may reveal a local mistake, a
missing capability, a harness defect, or an unreliable evaluator. Only a genuine capability
gap should trigger training. The loop converts diagnosed gaps into curriculum material,
updates the model through continual post-training, evaluates the resulting capability,
and then retries the original task. In an implementation, curriculum practice must remain
separate from the protected domain and retention tests used to accept an update.

\input{tex/report/synthesis/automated_continual_post_training_loop_figure}

Self-benchmarking creates the main reliability problem. If one model generates the tasks,
answers them, judges them, and trains on them, the loop can reward its own misconceptions,
leak evaluation items into training, or construct an increasingly easy curriculum. A
credible loop therefore separates training tasks from held-out tests, grounds verification
in an environment or independent judge, records data provenance, evaluates regression on
earlier capabilities, and retains a human gate or rollback path when evidence is uncertain.
Self-distillation has shown that sequential skills can be acquired with less forgetting
than ordinary SFT, but this is evidence for one update mechanism, not yet for a fully
autonomous improvement loop \citep{shenfeld2026selfdistillation}.

This distinction also separates two ambitions. Automated domain adaptation is relatively
well anchored: a human supplies the target domain, source material, task constraints, and
often verifiable tests. Recursive self-improvement is stronger because the system must also
improve its gap detector, curriculum builder, and verifier without losing contact with an
external objective. Progress should therefore mean improvement on protected, independently
measured tasks with capability retention---not merely higher reward on the system's own
generated benchmark.

The earlier methods occupy different roles inside this loop. SFT clones verified examples;
RL searches against outcome evidence; critics and semantic feedback locate errors; OPD
transfers a correction on student-visited states; and MOPD merges specialized teachers.
None closes the loop alone. The position of this paper is that long-term agent improvement
requires their orchestration around trustworthy gap detection, hierarchical feedback,
continual evaluation, and retention.
