\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{microtype}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{xcolor}
\usepackage[colorlinks=true,allcolors=blue!55!black]{hyperref}

\geometry{margin=1in}
\setlist{nosep}
\title{From Sparse Outcomes to Dense Process Feedback:\\
LLM Post-Training for Reasoning and Long-Horizon Agents}
\author{Yunhzou}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This position paper studies how supervision and credit assignment shape large language
model post-training, from the Transformer architecture and outcome-reward reinforcement
learning to process critics, on-policy distillation, and continual self-distillation. The
central position, evidence, qualifications, and conclusions will be developed after the
five selected works have been read in depth.
\end{abstract}

\tableofcontents

\section{Scope and guiding questions}
\subsection{Research area}
Define LLM post-training for reasoning and long-horizon agents, emphasizing the problem
of assigning useful learning signals to decisions within autoregressive trajectories.

\subsection{Guiding questions}
\begin{enumerate}
  \item What behavior can SFT, outcome-reward RL, process supervision, and distillation
        each teach or preserve?
  \item When does terminal reward become insufficient for long-horizon credit assignment?
  \item How should a process reward model or critic evaluate intermediate states and actions?
  \item Why does training on the student's state distribution change generalization and recovery?
  \item Can on-policy self-distillation support continual improvement without catastrophic forgetting?
\end{enumerate}

\subsection{Selected works}
The five primary works are \emph{Attention Is All You Need}
\citep{vaswani2017attention}, \emph{DeepSeek-R1} \citep{guo2025deepseekr1},
\emph{On-Policy Distillation} \citep{lu2025onpolicydistillation},
\emph{AgentPRM} \citep{xi2025agentprm}, and \emph{Self-Distillation Enables
Continual Learning} \citep{shenfeld2026selfdistillation}. Technical reports and
research articles count as primary works for this working selection.

\subsection{Working position}
State the paper's eventual thesis here. A candidate structure is that increasingly
long-horizon agent training requires a deliberate combination of outcome signals,
localized process feedback, and on-policy distribution matching rather than a single
universal post-training objective.

\section{Paper reviews}
\input{tex/papers/paper-01}
\input{tex/papers/paper-02}
\input{tex/papers/paper-03}
\input{tex/papers/paper-04}
\input{tex/papers/paper-05}

\section{Cross-paper synthesis}
\subsection{Supervision and credit-assignment spectrum}
Compare next-token supervision, terminal outcome rewards, step-level reward/value
estimates, teacher token distributions, and privileged-context self-distillation.

\subsection{Training distributions and support}
Compare teacher-forced, offline, and on-policy state distributions. Analyze exposure
bias, exploration, support mismatch, and policy lag.

\subsection{Long-horizon agents}
Compare critic-free grouped policy optimization with process reward models and learned
value estimates. Separate step correctness, progress, promise of eventual success, and
terminal task completion.

\subsection{Continual learning and retention}
Analyze whether distillation can transfer or recover skills across post-training stages
without collapsing prior capabilities. Separate empirical evidence from hypotheses about
automated continual post-training.

\subsection{Supporting systems case studies}
Use the GLM-5 technical report and GLM-5.2 research article as examples of asynchronous
agent RL and cross-stage on-policy distillation \citep{glm5team2026,glm52team2026}.
Use Composer 2.5 as an example of combining trajectory-level RL with localized textual
feedback and an on-policy KL loss at a problematic turn \citep{cursor2026composer25}.
Use \emph{Let's Verify Step by Step} as foundational context for process versus outcome
supervision and PRM evaluation \citep{lightman2023verify}.

\subsection{Evidence and open questions}
Distinguish claims supported by controlled experiments from system-level correlations,
implementation reports, and personal hypotheses. Record incompatible evaluation and
compute-accounting assumptions before comparing results.

\section{Connection to experiments}
Relate the literature to the post-training experiments in this repository, including
which findings reproduced, which assumptions did not transfer, and what should be tested
next. Planned experiments include outcome-only GRPO, process/value critics, on-policy
distillation, localized privileged feedback, and cross-stage retention tests.

\section{Conclusion}
Summarize the current understanding of LLM post-training and the most important unresolved
research questions.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
