\subsection{GRPO background and algorithm overview}

Group Relative Policy Optimization trains a language model from scalar rewards assigned
to complete sampled responses. For each prompt, it generates several responses, compares
their rewards, and uses the relative result to change the probabilities of the sampled
token actions. The central training unit is therefore a group: one prompt and multiple
sampled response trajectories.

Let $q$ denote a prompt and let
$o_i=(o_{i,1},\ldots,o_{i,T_i})$ denote sampled response $i$. Using equal trajectory
weighting, the complete loss studied in this section is
\begin{equation}
  \begin{aligned}
  \mathcal{L}_{\mathrm{GRPO}}(\theta)
  =-\frac{1}{G}\sum_{i=1}^{G}\frac{1}{T_i}\sum_{t=1}^{T_i}
  \Big[&\min\!\left(
      \rho_{i,t}(\theta)A_i,
      \operatorname{clip}(\rho_{i,t}(\theta),1-\epsilon,1+\epsilon)A_i
    \right) \\
    &-\beta\big(\exp(d_{i,t})-d_{i,t}-1\big)\Big].
  \end{aligned}
  \label{eq:grpo-full-loss}
\end{equation}
Its three constructed quantities are
\begin{align}
  A_i
  &=\frac{r_i-\mu_q}{\sigma_q+\epsilon_A},
  &\mu_q&=\frac{1}{G}\sum_{j=1}^{G}r_j,
  &\sigma_q^2&=\frac{1}{G}\sum_{j=1}^{G}(r_j-\mu_q)^2,
  \label{eq:grpo-overview-advantage}\\
  \rho_{i,t}(\theta)
  &=\frac{\pi_\theta(o_{i,t}\mid q,o_{i,<t})}
          {\pi_{\mathrm{old}}(o_{i,t}\mid q,o_{i,<t})},
  \label{eq:grpo-overview-ratio}\\
  d_{i,t}
  &=\log\pi_{\mathrm{ref}}(o_{i,t}\mid q,o_{i,<t})
    -\log\pi_\theta(o_{i,t}\mid q,o_{i,<t}).
  \label{eq:grpo-overview-kl-difference}
\end{align}
Here $G$ is the number of responses sampled for one prompt, $T_i$ is the number of
trainable tokens in response $i$, and $r_i$ is its scalar reward. The advantage $A_i$
compares the complete response with the other responses in its group and is reused for
all of its tokens. The ratio $\rho_{i,t}$ compares the current trainable policy
$\pi_\theta$ with the frozen behavior policy $\pi_{\mathrm{old}}$ that sampled the token.
The clipping width $\epsilon$ limits how much the same sampled token can encourage one
update, while $\epsilon_A$ prevents division by zero when normalizing rewards. The
quantity $d_{i,t}$ constructs a sampled KL cost relative to a separate frozen reference
policy $\pi_{\mathrm{ref}}$, and $\beta$ sets the strength of that cost. The leading minus
sign converts a quantity that should be maximized into a loss minimized by the optimizer.

Equation~\ref{eq:grpo-full-loss} is the map for the rest of the section: first construct
$A_i$, then derive $\rho_{i,t}$, and finally explain clipping and reference-policy KL.
Operationally, GRPO performs the following high-level loop:
\begin{enumerate}
  \item freeze the current model and use that snapshot to sample $G$ responses to the
        same prompt;
  \item assign one reward to each complete response;
  \item compare rewards within the group to obtain one advantage per response;
  \item replay the sampled tokens through the trainable policy;
  \item increase the probabilities of actions from positive-advantage responses and
        decrease those from negative-advantage responses;
  \item clip large changes relative to the frozen sampling snapshot and penalize drift
        from a separate frozen reference model; and
  \item update the model and generate a new rollout group.
\end{enumerate}

The next modules unpack these steps in exactly this order. The initial explanation assumes
a synchronous learner and a simple response containing only model-generated tokens.
Rollout buffers, asynchronous workers, and agentic trace construction are introduced only
after the basic update is established. DeepSeek-R1 uses GRPO during its reinforcement-
learning stages \citep{guo2025deepseekr1}; this attribution identifies the source studied
here but is not needed to understand the algorithm.
