\subsection{Why GRPO clips movement and retains a reference policy}
\label{subsec:grpo-clipped-objective}

The local reward surrogate is most trustworthy near the policy that generated its data.
After an optimizer step, however, the trainable policy changes while the stored actions do
not. Reusing the buffer may be computationally attractive because generation is expensive,
but each additional step asks increasingly stale samples to justify further movement.

For sampled token $o_{i,t}$, the learner measures that mismatch by recomputing its
log-probability and comparing it with the stored behavior-policy log-probability:
\begin{equation}
  \rho_{i,t}(\theta)
  =\exp\!\left(
      \log\pi_\theta(o_{i,t}\mid q,o_{i,<t})
      -\log\pi_{\mathrm{old}}(o_{i,t}\mid q,o_{i,<t})
    \right).
  \label{eq:grpo-token-ratio}
\end{equation}
A ratio near one says the sample still describes the current policy reasonably well. A
large departure says that the learner is trying to draw a strong conclusion about a
policy that did not generate this data. Clipping refuses to grant additional reward for
moving farther in the already encouraged direction:
\begin{equation}
  S_{i,t}(\theta)
  =\min\!\left(
      \rho_{i,t}A_i,
      \operatorname{clip}(\rho_{i,t},1-\epsilon,1+\epsilon)A_i
    \right).
  \label{eq:grpo-clipped-surrogate}
\end{equation}

Suppose $\epsilon=0.2$. For positive advantage $A_i=1.5$ and ratio $\rho=1.5$, the raw
credit is $2.25$, but the clipped candidate is $1.2(1.5)=1.8$; the minimum selects $1.8$.
The old sample provides no further reward for making its action more than $20\%$ likelier.
For negative advantage $A_i=-1$ and ratio $\rho=0.5$, the raw term is $-0.5$ and the
clipped candidate is $-0.8$; the minimum selects $-0.8$. Further suppression earns no
additional improvement. Clipping is therefore a limit on what one stale batch may justify,
not a new reward signal.

Clipping only compares the policy with the snapshot that generated the current buffer.
That snapshot is refreshed, so many individually small updates can still move the model
far from its starting behavior. This matters because the reward measures only a narrow
proxy: a math checker does not measure conversational quality, factual breadth, or every
way the policy might exploit the checker. GRPO therefore keeps a separate frozen reference
policy as a long-term anchor.

Using the sampled-action KL estimator in this objective, define
\begin{equation}
  d_{i,t}
  =\log\pi_{\mathrm{ref}}(o_{i,t}\mid q,o_{i,<t})
   -\log\pi_\theta(o_{i,t}\mid q,o_{i,<t}),
  \qquad
  \widehat D_{i,t}=\exp(d_{i,t})-d_{i,t}-1.
  \label{eq:grpo-sampled-kl}
\end{equation}
The token-level score combines the bounded reward incentive with the reference cost:
\begin{equation}
  J_{i,t}^{\mathrm{GRPO}}(\theta)
  =S_{i,t}(\theta)-\beta\widehat D_{i,t}.
  \label{eq:grpo-token-objective}
\end{equation}
Higher reward credit improves this score; greater reference drift reduces it. Averaging
over tokens and sibling responses, then negating because software minimizes losses,
reassembles the full objective:
\begin{equation}
  \mathcal{L}_{\mathrm{GRPO}}(\theta)
  =-\frac{1}{G}\sum_{i=1}^{G}\frac{1}{T_i}\sum_{t=1}^{T_i}
    J_{i,t}^{\mathrm{GRPO}}(\theta)
  =-\frac{1}{G}\sum_{i=1}^{G}\frac{1}{T_i}\sum_{t=1}^{T_i}
    \left(S_{i,t}(\theta)-\beta\widehat D_{i,t}\right).
  \label{eq:grpo-reassembled-loss}
\end{equation}
Equation~\ref{eq:grpo-reassembled-loss} is exactly
Equation~\ref{eq:grpo-full-loss} written with the shorthand built in this subsection.
Only now is the historical comparison useful: the clipped ratio construction was
introduced by Proximal Policy Optimization, while GRPO replaces its learned value
baseline with the sibling-group estimate \citep{schulman2017ppo}.
