\subsection{From one scalar judgment to a policy gradient}
\label{subsec:grpo-gradient-direction}

After sibling comparison, a sampled response may have advantage $+1$ or $-1$. The checker
still does not say which replacement tokens would be better, and its score cannot be
differentiated. The only honest update is therefore to reweight the behavior the model
actually produced: make a positive-advantage response more likely and a
negative-advantage response less likely.

Consider only two possible responses. The behavior policy gives each probability $0.5$;
one has advantage $+1$ and the other $-1$. If a new policy moves their probabilities to
$0.7$ and $0.3$, it has moved in the desired direction because
\begin{equation}
  0.7(+1)+0.3(-1)=0.4>0.
  \label{eq:grpo-parent-objective-example}
\end{equation}
The verbal goal is therefore: adjust the policy so that a response drawn from it tends to
have positive advantage. Mathematically, that goal is the expected-advantage objective
\begin{equation}
  J_{\mathrm{adv}}(\theta)
  =\mathbb{E}_{o\sim\pi_\theta(\cdot\mid q)}
    \left[A_{\mathrm{old}}(q,o)\right].
  \label{eq:grpo-parent-expected-advantage}
\end{equation}
The advantages are held fixed during this optimizer step. They judge the samples; the
trainable object is the probability that the policy assigns to those samples.

Why does a log-probability gradient appear? Write the expectation as a probability-
weighted sum and differentiate it. The identity
$\nabla_\theta\pi_\theta=\pi_\theta\nabla_\theta\log\pi_\theta$ turns the result back into
an expectation that can be estimated with sampled responses:
\begin{equation}
  \nabla_\theta J_{\mathrm{adv}}
  =\mathbb{E}_{o\sim\pi_\theta}
    \left[A_{\mathrm{old}}(q,o)\nabla_\theta\log\pi_\theta(o\mid q)\right].
  \label{eq:grpo-parent-policy-gradient}
\end{equation}
This formula is not an arbitrary loss design. It is the derivative of ``put more
probability on responses with positive advantage.'' The log appears because it provides a
sample-based way to differentiate a probability distribution when the reward itself is a
black box.

An autoregressive model assigns a response probability by multiplying its conditional
token probabilities. Taking a logarithm turns that product into a sum:
\begin{equation}
  \log\pi_\theta(o_i\mid q)
  =\sum_{t=1}^{T_i}\log\pi_\theta(o_{i,t}\mid q,o_{i,<t}).
  \label{eq:grpo-response-logprob-decomposition}
\end{equation}
Consequently, one response-level advantage weights the gradients of every sampled token
decision. For a positive-advantage response `5 <eos>', the model reinforces both emitting
`5' from the prompt and stopping after the prefix containing `5'. For a
negative-advantage response `4 <eos>', both sampled decisions are suppressed. This is
coarse outcome-level credit assignment: the final judgment does not identify which token
deserved credit.

The principle is now clear, but the current buffer creates a practical mismatch. The
desired expectation concerns the trainable policy $\pi_\theta$, while the stored answers
came from the frozen behavior policy $\pi_{\mathrm{old}}$. The next subsection connects
the desired update to those stored samples.
