\subsection{Compute, storage, and algorithmic limitations}

GRPO adds three major costs beyond an ordinary language-model training step. First,
grouped rollout generation requires $G$ autoregressive responses per prompt and is
sequential across the tokens of each response. Second, reward computation may require
verifiers, reward models, tools, or external environments. Third, optimization requires a
current-policy forward and backward pass over the stored response tokens, plus forward
evaluation of the reference policy if reference log-probabilities were not precomputed.

For $M$ stored response tokens, the rollout buffer requires $O(M)$ token-level storage for
action IDs, old log-probabilities, masks, and optionally reference log-probabilities. Rewards
and advantages require only one scalar per trajectory before broadcasting. The trainable
policy carries parameters, gradients, and optimizer states. The reference policy is frozen
and needs no gradients or optimizer states. A separate old-policy model is unnecessary
when the behavior log-probabilities of the sampled actions have been stored exactly.

The basic GRPO update has several limitations:
\begin{itemize}
  \item one outcome advantage is broadcast across every trainable action, so it cannot
        identify which reasoning step caused success or failure;
  \item prompts whose sampled responses all receive the same reward provide no relative
        advantage signal;
  \item larger groups improve the reward baseline but increase rollout cost;
  \item the token-reduction rule determines whether long trajectories contribute more
        total gradient than short trajectories;
  \item buffer reuse and asynchronous lag make advantages and sampled prefixes stale; and
  \item imperfect verifiers can reward formatting shortcuts or other reward-hacking
        behavior rather than the intended reasoning process.
\end{itemize}

These limitations do not change the basic GRPO derivation. They motivate later methods
that change the quality or granularity of the training signal, and systems that preserve
the validity of the trajectory supplied to the update.
