\subsection{Complete GRPO training loop}

One synchronous GRPO iteration proceeds chronologically as follows:
\begin{enumerate}
  \item Copy the current policy version into the behavior policy
        $\pi_{\mathrm{old}}$.
  \item For every prompt, sample $G$ response trajectories from
        $\pi_{\mathrm{old}}$ without recording gradients.
  \item Store the exact sampled token IDs and their behavior-policy log-probabilities.
  \item Score every complete trajectory and compute its group-relative advantage using
        Equation~\ref{eq:grpo-group-advantage}.
  \item Replay the stored trajectories through $\pi_\theta$ with teacher forcing and
        gather the current log-probabilities of the sampled actions.
  \item Compute token ratios, clipped advantage terms, and reference-policy KL penalties
        using Equations~\ref{eq:grpo-token-ratio}--\ref{eq:grpo-token-objective}.
  \item Apply the response-token mask and reduce all valid token terms to one scalar loss.
  \item Run backpropagation, optionally clip the parameter-gradient norm, and take an
        optimizer step.
  \item If the configuration reuses the buffer, repeat Steps 5--8 for the selected
        minibatches or epochs. Otherwise discard it immediately.
  \item Discard the completed buffer, refresh the behavior policy from the updated
        $\pi_\theta$, and generate fresh trajectories.
\end{enumerate}

The rollout phase and optimization phase perform different computations. Rollout samples
one token at a time and stores the realized actions. Optimization does not sample a second
response; it teacher-forces the stored response so that the probabilities of those exact
actions can be differentiated. Changing tokenization, deleting an action, or substituting
a reconstructed response would change the trajectory being optimized.
