\subsection{On-policy self-distillation from privileged feedback}

In OPSD, the ``S'' means \emph{self}: the policy supplies its own teacher signal under a
better context. At a selected student prefix $s_t$, the ordinary policy produces
$P_t=\pi_\theta(\cdot\mid s_t)$. A teacher branch of the same policy additionally sees a
validated hint $h_t$ and produces $Q_t=\pi_\theta(\cdot\mid s_t,h_t)$. Conditioning
on the hint steers $Q_t$ toward a more appropriate next action. The teacher branch is
treated as fixed for the update, and minimizing $D_{\mathrm{KL}}(P_t\|Q_t)$ pulls the
ordinary, no-hint distribution toward that correction. The prefixes are on-policy because
the ordinary student visited them; the target is self-distilled because the same model
supplies $Q_t$ under privileged context
\citep{shenfeld2026selfdistillation,cursor2026composer25}.
The intuition is direct: if the model behaves correctly when reminded, distill the
reminded distribution so that it can behave that way without the reminder. The
continual-learning study stabilizes this self-teacher with exponential-moving-average
weights by default; it does not introduce an external expert.

``Privileged'' should not mean answer leakage. A valid hint supplies feedback or a
constraint that the trainer can legitimately know---for example, the tools actually
available at that turn, a grounded diagnosis of a local error, or a required communication
rule. It should not expose a held-out answer, hidden tests, or future environment results.
Otherwise the teacher distribution is improved by information unavailable to the task,
and the resulting benchmark or capability claim is contaminated.

Composer 2.5 applies this construction to localized mistakes inside long RL rollouts. If
the policy calls an unavailable tool, a short reminder listing the available tools is added
only to the teacher context. The hint lowers the teacher probability of the invalid call
and raises valid alternatives; a KL loss at that turn moves the original-context policy in
the same direction, while the complete rollout retains its RL objective
\citep{cursor2026composer25}. At inference the hint is absent: its behavioral effect has
been compiled into the model parameters.

A distinct extension would train the model to generate and use grounded diagnoses itself,
using sequences such as \emph{state $\rightarrow$ feedback $\rightarrow$ corrected
action}. This may create a reusable inference-time reflection skill, but it is not what
OPSD itself does. OPSD internalizes the distributional effect of supplied feedback without
requiring the model to reproduce the hint.
