\subsection{Initialization, efficiency, and limits}

OPD is most useful when the student already has the relevant capability in its sampling
support and a trustworthy teacher is available. A teacher forward pass supplies feedback
at every token, partial rollouts can be trained without waiting for a terminal verifier,
and no separate reward model is required. In the reported reasoning experiment, OPD raised
AIME'24 accuracy from roughly $60\%$ to $70\%$ using about 77,000 prompts after the SFT
initialization; the article's compute comparisons favor OPD, but depend on its models,
systems, and accounting assumptions \citep{lu2025onpolicydistillation}.

OPD needs local policy overlap, not identical models. The teacher can strongly penalize a
bad student token, but dense feedback does not create missing support. If the student is
far away, its rollouts mostly visit poor prefixes and may never expose the teacher's useful
strategy. The update then says what to suppress without demonstrating the missing path;
SFT, mid-training, or an easier teacher may first be needed to bridge the policies. Recent
experiments likewise identify large student--teacher mismatch as a case where the guiding
signal can become misaligned with task correctness \citep{wang2026demystifyingopd}.

Long horizons compound this mismatch. A small probability of leaving teacher-like states
at each step makes late student prefixes increasingly unlike prefixes the teacher would
normally produce. Later tokens are not automatically given less loss weight---the teacher
still scores every actual prefix---but its local correction may become less useful far off
its normal trajectory distribution. Alternating sample, update, and resample rounds moves
the visited distribution gradually and partially mitigates the problem, provided the
initial overlap supplies a learnable bridge. OPD also inherits teacher errors and may
reduce diversity through reverse-KL mode seeking; RL is still needed when success must be
discovered or judged externally rather than transferred from a teacher.

The empirical claim must remain narrower than the intuition. Privileged-context
self-distillation has degraded thinking models most at long rollout budgets, apparently by
suppressing useful reconsideration at high-entropy branch points; the same study reports
that ordinary OPD can still help \citep{kaur2026rethinkingopsd}. Thus, horizon length alone
does not invalidate OPD. Teacher distance, privileged-context strength, and where the
distillation loss is applied jointly determine whether later feedback remains useful.
