\subsection{From a conversation to supervised tokens}

The model does not receive a list of abstract roles. A chat template serializes every
turn, including role markers and end-of-turn tokens, into one causal sequence. For a
minimal prompt and answer, an abstract serialization is
\begin{verbatim}
tokens:  [BOS, SYSTEM, S1, USER, Q1, Q2, ASSISTANT, A1, A2, EOS]
inputs:  [BOS, SYSTEM, S1, USER, Q1, Q2, ASSISTANT, A1, A2]
targets: [SYSTEM, S1, USER, Q1, Q2, ASSISTANT, A1, A2, EOS]
score:   [   no,   no,   no, no, no,        no, yes, yes, yes]
\end{verbatim}
The logit after \texttt{ASSISTANT} predicts the first response token \texttt{A1}; the
logit after \texttt{A2} predicts \texttt{EOS}. This alignment is easy to shift by one
position accidentally. Some libraries shift labels inside the model and others expect
already shifted targets, but the shift must happen exactly once.

System and user tokens remain in the input because the response conditions on them. A
response-only mask merely removes their direct next-token targets. Gradients from an
assistant-token loss can still flow backward through prompt representations that the
assistant position attended to.

The same ownership rule extends to agent trajectories. Assistant actions are normally
supervised; tool and environment responses remain context but are not relabeled as
assistant actions. An auxiliary objective may deliberately predict selected observation
tokens. ECHO is one example of this separation \citep{shrivastava2026echo}. That is an
additional environment-modeling loss, not ordinary response-only SFT, and arbitrary
private, noisy, or transient tool output may be a poor target.
