\subsection{Behavior cloning and response masking}

SFT presents a condition and an expert continuation, then increases the continuation's
probability. A chat template first serializes system, user, assistant, and tool turns into
one causal token sequence. Prompt and tool tokens remain visible as context, but a
response-only mask normally assigns direct loss only to assistant tokens. If $m_{b,t}$
marks those targets and $M=\sum_{b,t}m_{b,t}$, the loss is
\begin{equation}
  \mathcal{L}_{\mathrm{SFT}}
  =-\frac{1}{M}\sum_{b,t}m_{b,t}
    \log p_\theta(y_{b,t+1}\mid y_{b,\leq t}).
  \label{eq:sft-loss}
\end{equation}
This is ordinary next-token cross-entropy with target selection, not a new model
architecture. $M$ counts supervised tokens rather than padded sequence length. Masking
does not avoid the forward computation for prompt context; it only decides which predicted
positions contribute directly to the scalar loss.

Token averaging creates a dataset-length imbalance. For example, if a batch contains one
20-token response and one 980-token response, their approximate weights in the gradient
are $2\%$ and $98\%$, even though the dataset contains one trajectory of each. A useful
direction represented by the short example can therefore be diluted by the long trace.
This is a statistical choice, not a padding problem: packing and length-grouped batching
improve utilization but do not change the weights. Per-example averaging, task-balanced
sampling, or explicit weights instead choose demonstrations or tasks, rather than tokens,
as the equal units.
