\subsection{What the demonstration distribution can teach}

SFT can teach instruction following, answer format, tool syntax, terminology, length, and
reasoning style when these behaviors are represented consistently in the demonstrations.
Its generalization is constrained by what varies in that dataset.

\paragraph{System prompts as conditions.}
Including a system prompt in the input does not guarantee that the learned behavior
remains controllable through that prompt. If exactly the same system message appears in
every example, it is a constant rather than an informative variable. The data provide no
contrast showing which behavior should change when the message changes, so repeated SFT
can absorb its instruction into the model parameters. There is no universal iteration at
which this happens, and a larger training run does not solve the identifiability problem.
Prompt control is better tested and taught by varying system instructions, including
contrasting instructions for comparable user inputs, sometimes omitting generic prompts,
and evaluating behavior when the prompt is changed or removed.

\paragraph{One demonstrated path.}
A demonstration selects one valid token path. Cross-entropy increases that path's local
probability even when other continuations would also be correct. Repetitive data can
therefore reduce diversity, reinforce template artifacts, and shift capability away from
domains absent from the mixture. Low training loss establishes imitation of the training
set, not generalization.

\paragraph{Length changes example weight.}
Let $\bar{g}_b$ be the mean gradient over the $M_b$ supervised tokens in trajectory $b$.
The token-mean gradient can be rewritten as
\begin{equation}
  g_{\mathrm{token}}
  =\sum_{b=1}^{B}\frac{M_b}{\sum_j M_j}\,\bar{g}_b.
\end{equation}
Thus trajectories with more supervised tokens have more influence on the update direction.
This is the expected maximum-likelihood objective when each next-token decision is the
unit of data, but it is not automatically the right task mixture. A verbose reasoning
trace may dominate a short tool call even if the two examples should receive equal task
weight.

For agent data, total context length and supervised length must be distinguished. A long
system prompt or tool response that is masked from the SFT loss increases attention and
activation cost but does not increase $M_b$. A long assistant reasoning or action span
does increase $M_b$. If selected environment tokens receive an ECHO-style auxiliary loss,
that objective has its own supervised-token count and coefficient; it should not silently
be folded into the assistant-token denominator.

\paragraph{Mixing short and long examples.}
Packing and length-grouped batching are systems optimizations: they reduce padding and
improve utilization, but they do not by themselves equalize trajectory weights. A useful
default is to mix lengths, then monitor both total processed length and supervised-token
length by task and data source. If long traces dominate unintentionally, the data loader
can balance task or length strata, the objective can use per-example normalization, or
supervised spans can be capped or explicitly weighted. Short- and long-context evaluation
should remain separate so that a single aggregate loss does not hide a regression.

Extreme context extension is often treated as a distinct empirical design problem rather
than an ordinary batching choice. LongSkywork used a specialized long-context SFT stage
after standard SFT \citep{zhao2024longskywork}. In contrast, the ProLong study found that
high-quality short data was important in its long-context training mixture and that short
instruction SFT alone could still yield strong long-context task performance
\citep{gao2024longcontext}. These results do not define one universal schedule; they show
that separating or mixing lengths should be justified by the target capability and
validated experimentally.

\paragraph{Teacher-forced prefixes.}
During SFT, the model predicts each response token after the correct demonstrated prefix.
During inference, described in Section~\ref{sec:transformer-inference}, each new token is
conditioned on the model's own earlier samples. An early mistake can lead to a prefix that
never appeared in the SFT data. This exposure-bias gap explains why dense demonstration
learning can produce strong initial behavior without directly teaching recovery from the
model's own errors.

SFT is therefore a strong initialization method when desired behavior can be demonstrated
clearly. It does not by itself identify every valid alternative, evaluate terminal
outcomes, or supply credit for actions taken on the model's own trajectory. Those are
separate problems for the later sections on reinforcement learning, critics, and
on-policy distillation.
