\subsection{Supervised fine-tuning}

Pretraining teaches a model to continue text from a broad corpus. Supervised fine-tuning
(SFT) changes the conditional behavior by training on demonstrations of desired responses.
For example, an instruction-tuning record may contain
\begin{center}
\begin{tabular}{ll}
  \toprule
  Role & Content \\
  \midrule
  System & You are a concise assistant. \\
  User & What is $2+2$? \\
  Assistant & 4 \\
  \bottomrule
\end{tabular}
\end{center}
The chat template converts these structured fields into one causal token sequence, for
example
\begin{center}
\texttt{[SYSTEM, You, are, concise, USER, What, is, 2+2, ASSISTANT, 4, EOS]}.
\end{center}
The system, user, and assistant delimiters are ordinary conditioning tokens from the
model's perspective, although their IDs are reserved by the tokenizer.

\subsubsection{Response-only next-token loss}

SFT commonly retains every formatted token as context but calculates loss only when the
next-token target belongs to the assistant. An abstract alignment is
\begin{center}
\begin{tabular}{lll}
  \toprule
  Input at position $t$ & Next-token target & Scored? \\
  \midrule
  \texttt{SYSTEM} & first system-content token & no \\
  final system token & \texttt{USER} & no \\
  final user token & \texttt{ASSISTANT} & no \\
  \texttt{ASSISTANT} & \texttt{4} & yes \\
  \texttt{4} & \texttt{EOS} & yes \\
  \bottomrule
\end{tabular}
\end{center}
If $m_{b,t}\in\{0,1\}$ is this ownership mask and $M=\sum_{b,t}m_{b,t}$, the objective is
\begin{equation}
  \mathcal{L}_{\mathrm{SFT}}
  =-\frac{1}{M}\sum_{b,t}m_{b,t}
    \log p_\theta(y_{b,t+1}\mid y_{b,\leq t}).
\end{equation}
At a selected position, cross-entropy increases the demonstrated token's logit and
decreases competing logits. If the model assigns probability 0.4 to the demonstrated
token, that position contributes $-\log(0.4)\approx0.916$ nats.

As in the Transformer training discussion, $1/M$ normalizes rather than avoids
computation. Masked prompt positions still pass through every layer. They influence the
assistant hidden states, and assistant-loss gradients can flow through their
representations. The mask only removes their direct next-token targets. This distinction
is essential in multi-turn agent data:
\begin{center}
\begin{tabular}{lcc}
  \toprule
  Segment & Present in context? & Policy loss? \\
  \midrule
  System and user tokens & yes & normally no \\
  Assistant tokens & yes & yes \\
  Tool/environment observations & yes & normally no policy loss \\
  Padding & no useful content & no \\
  \bottomrule
\end{tabular}
\end{center}
Training on tool output as an assistant target would teach the policy to imitate its
environment rather than to act after observing it. This does not imply that observations
can never be supervised. ECHO, for example, keeps the roles separate while adding an
auxiliary next-token loss on selected terminal-response tokens alongside the policy loss
on assistant actions \citep{shrivastava2026echo}. Such a loss is a deliberate
environment-modeling objective, not ordinary response-only SFT, and noisy, private,
transient, or irrelevant tool outputs may be poor prediction targets.

\subsubsection{SFT algorithm}

A response-only SFT update is:
\begin{enumerate}
  \item sample demonstrations and apply the exact deployment chat template;
  \item tokenize the complete conversations and record token ownership;
  \item construct next-token labels, replacing non-assistant targets with an ignore index;
  \item run one teacher-forced causal forward pass over each complete sequence;
  \item average cross-entropy over selected assistant targets;
  \item backpropagate and update the model parameters.
\end{enumerate}
The InstructGPT pipeline is a prominent example: demonstrations first produced an SFT
policy, which was then used to initialize later preference-based RL
\citep{ouyang2022instructgpt}. This ordering reflects a general role for SFT: behavior
cloning places useful responses inside the policy's high-probability region before a
weaker or sparser feedback signal attempts further optimization.

Teacher forcing makes SFT dense and stable. All demonstrated response positions can be
scored in parallel because the complete response is known and the causal mask prevents
future leakage. However, deployment conditions on the model's sampled history. If it
makes an early mistake, it may reach a prefix absent from the demonstration data. This
train--inference state-distribution difference is exposure bias.

\subsubsection{What SFT controls and what it cannot specify}

The data distribution determines which behavior is cloned. Demonstrations can teach
format, instruction following, tool syntax, terminology, length, and reasoning style.
Masking prompt targets keeps long inputs from dominating the objective, but it does not
make behavior invariant to prompt wording. If every demonstration uses the same system
prompt, that prompt is a constant condition. Its behavior may be absorbed into the policy
rather than remain controllable when the prompt is removed or changed.

SFT also observes only selected response paths. A demonstration containing one correct
answer does not label every alternative answer as semantically wrong, yet token
cross-entropy moves probability toward the demonstrated next token and away from its
competitors. Narrow or repetitive data can consequently reduce diversity, solidify
template artifacts, and cause capability regression outside the fine-tuning domain.

Two averaging conventions further change the effective data weighting. A global token
mean gives each response token equal weight, so long responses contribute more total
gradient. Averaging within each example before averaging the batch gives each
demonstration equal total weight. This choice becomes consequential when short tool calls
and thousand-token reasoning traces occur in the same corpus.

SFT is therefore best viewed as dense off-policy behavior cloning. It is efficient when
high-quality target behavior can be demonstrated, but it does not directly train recovery
on the learner's own mistakes or optimize a non-differentiable outcome. These limitations
motivate the later comparison with outcome-reward RL, process critics, and on-policy
distillation.
