\section{Proposed method: hinted SFT}

Hinted SFT separates \emph{discovering the correction} from \emph{acting under the
correction}. The first stage trains the missing intermediate directly. The second stage
ensures that later actions actually depend on that intermediate. Figure~\ref{fig:hinted-sft}
shows the complete training and deployment paths.

\begin{figure}[ht]
  \centering
  \begin{tikzpicture}[node distance=5mm,scale=0.80,transform shape]
    \node[font=\bfseries,text=ProposalInk,anchor=east] (trainlabel) at (-6.2,0) {Training};
    \node[proposalbox,text width=29mm] (trainstate) at (-4.2,0)
      {Visible state $x$\\history + evidence};
    \node[supportedbox,text width=29mm,right=of trainstate] (reviewer)
      {Reviewer creates\\verified hint $h^\star$};
    \node[actionbox,text width=29mm,right=of reviewer] (hintloss)
      {SFT on\\hint tokens};
    \node[supportedbox,text width=29mm,right=of hintloss] (context)
      {Append $h^\star$\\to live context};
    \node[actionbox,text width=29mm,right=of context] (contloss)
      {Train or evaluate\\the continuation};
    \draw[proposalarrow] (trainstate) -- (reviewer);
    \draw[proposalarrow] (reviewer) -- (hintloss);
    \draw[proposalarrow] (hintloss) -- (context);
    \draw[proposalarrow] (context) -- (contloss);

    \node[font=\bfseries,text=ProposalInk,anchor=east] (deploylabel) at (-6.2,-2.0) {Deployment};
    \node[proposalbox,text width=29mm] (deploystate) at (-4.2,-2.0)
      {Visible state $x$\\history + evidence};
    \node[actionbox,text width=29mm,right=of deploystate] (generate)
      {Model generates\\self-hint $\hat h$};
    \node[supportedbox,text width=29mm,right=of generate] (gate)
      {Grounding gate\\accept / retrieve / abstain};
    \node[supportedbox,text width=29mm,right=of gate] (livecontext)
      {Append $\hat h$\\to live context};
    \node[actionbox,text width=29mm,right=of livecontext] (act)
      {Model acts\\under $\hat h$};
    \draw[proposalarrow] (deploystate) -- (generate);
    \draw[proposalarrow] (generate) -- (gate);
    \draw[proposalarrow] (gate) -- (livecontext);
    \draw[proposalarrow] (livecontext) -- (act);

    \node[riskbox,text width=40mm,below=6mm of gate] (verify)
      {Environment verifies outcome\\and stores grounded failures};
    \draw[proposalarrow] (act.south) |- (verify.east);
    \draw[proposalarrow,dashed] (verify.west) -| node[below left,font=\scriptsize]{next training round} (generate.south);
  \end{tikzpicture}
  \caption{\textbf{Hinted SFT makes the correction an explicit causal intermediate.}
  Training supervises the hint that the model must later generate. Deployment conditions
  the continuation on the model's own accepted hint, while the environment supplies
  evidence for verification and subsequent data generation.}
  \label{fig:hinted-sft}
\end{figure}

\subsection{Training unit}

Each example contains four aligned objects:
\[
  (x,\ e,\ h^\star,\ y^\star),
\]
where $x$ is the trajectory prefix, $e$ is deployment-visible evidence, $h^\star$ is a
short verified diagnosis or constraint, and $y^\star$ is an optional corrected
continuation. The serialized trace can use an explicit internal channel:
\[
  [x;\ e;\ \texttt{<reflection>}\ h^\star\ \texttt{</reflection>};\ y^\star].
\]
The reflection channel need not be shown to the user, but it must be present in the
model's autoregressive context before $y^\star$. This ordering is essential: a diagnosis
generated after the action is a rationalization, not a cause of that action.

The core hint loss is ordinary response-masked SFT:
\[
  \mathcal L_{\mathrm{hint}}
  =-\frac{1}{|h^\star|}\sum_j
    \log \pi_\theta(h^\star_j\mid x,e,h^\star_{<j}).
\]
System instructions, user input, and environment evidence remain context; the verified
hint tokens are the primary targets. Three continuation variants should be compared:

\begin{enumerate}
  \item \textbf{Hint only (HSFT-H):} train only $\mathcal L_{\mathrm{hint}}$ and allow the
  existing policy to respond to the generated hint.
  \item \textbf{Hint plus continuation (HSFT-HC):} add SFT loss on $y^\star$. This is
  stronger behavior cloning but may increase forgetting.
  \item \textbf{Hint plus on-policy continuation (HSFT-H+OPD/RL):} train the hint with SFT,
  then sample continuations under the generated hint and improve them with dense
  distillation or verified reward. This tests whether explicit hint generation and OPSD
  are complementary rather than mutually exclusive.
\end{enumerate}

\subsection{On-policy self-hint refinement}

Teacher-forcing $h^\star$ during training creates its own train--test mismatch: deployment
uses $\hat h\sim\pi_\theta(\cdot\mid x,e)$. The proposal therefore uses three stages.

\begin{enumerate}
  \item \textbf{Bootstrap:} SFT on verified reviewer hints.
  \item \textbf{Self-hint rollout:} sample $\hat h$ from the current model before acting.
  A verifier checks whether every factual claim in $\hat h$ is supported by $x$, $e$, or
  an allowed retrieval.
  \item \textbf{Iterative correction:} retain successful self-hints; relabel rejected
  hints with a grounded correction or an abstention target, then retrain. Continuations
  are always generated with the accepted self-hint in context.
\end{enumerate}

This loop prevents the proposal from hiding the original problem behind a new privileged
hint generator. The deployed model must demonstrate that it can generate the intermediate
from available evidence, not merely consume it.
