\section{Risks, scope, and expected contribution}

\paragraph{Rationalization instead of reasoning.}
A model may generate a plausible explanation for an action it had already selected
internally. The intervention test is structural: generate the hint first, alter or remove
it before decoding the action, and measure whether the action distribution changes
appropriately.

\paragraph{Hint hallucination.}
Explicit text can amplify confidence in a false premise. Hints should cite evidence spans,
tool outputs, or retrieved artifacts; a grounding gate should reject unsupported claims.
The target for unavailable evidence is retrieval or abstention, not a fabricated
diagnosis.

\paragraph{Reflection overhead and collapse.}
Always generating long reflections increases inference cost and can collapse into generic
phrases. Hints should be short, localized, and triggered by a learned gate. A no-hint
action remains available when confidence is high and no correction is needed.

\paragraph{Exposure bias and compounding errors.}
Training only on reviewer hints leaves the continuation unprepared for imperfect
self-hints. The staged self-hint rollout is therefore part of the method rather than an
optional implementation detail. Robustness should be measured under deliberately
perturbed but plausible hints.

\paragraph{Where OPSD may remain preferable.}
OPSD may be more efficient when the desired change is stylistic, when a hint merely
reweights already obvious alternatives, or when printing an intermediate adds no useful
control surface. Dense KL feedback can also reduce forgetting relative to strong
off-policy SFT. The proposal predicts a benefit specifically when an agent should
\emph{discover, expose, and reuse} a diagnosis during a long trajectory.

\paragraph{Expected contribution.}
The intended contribution is not a new loss function. It is a training decomposition and
evaluation protocol:

\begin{enumerate}
  \item define whether privileged feedback is supported by deployment-visible evidence;
  \item train the model to generate the missing, grounded intermediate when it is
  inferable;
  \item place that intermediate in the live context before the corrected action;
  \item verify causal use through counterfactual, removal, and hidden-evidence tests; and
  \item combine explicit semantic correction with OPD or RL only after the intermediate is
  available to the acting policy.
\end{enumerate}

If successful, hinted SFT would provide a practical bridge between textual feedback and
long-horizon agent learning. Instead of silently compiling every privileged intervention
into model weights, the model learns a reusable operation: inspect available evidence,
state the correction, and let that correction guide the next part of the trajectory.
