\section{Scope and guiding questions}

\subsection{Research area}
Define LLM post-training for reasoning and long-horizon agents, emphasizing the problem
of assigning useful learning signals to decisions within autoregressive trajectories.

\subsection{Guiding questions}
\begin{enumerate}
  \item What behavior can SFT, outcome-reward RL, process supervision, and distillation
        each teach or preserve?
  \item When does terminal reward become insufficient for long-horizon credit assignment?
  \item How should a process reward model or critic evaluate intermediate states and actions?
  \item Why does training on the student's state distribution change generalization and recovery?
  \item Can on-policy self-distillation support continual improvement without catastrophic forgetting?
\end{enumerate}

\subsection{Core source corpus}
The working corpus contains ten central sources:
\begin{enumerate}
  \item \emph{Attention Is All You Need} \citep{vaswani2017attention};
  \item \emph{DeepSeek-R1} \citep{guo2025deepseekr1};
  \item \emph{Polar: Agentic RL on Any Harness at Scale} \citep{xu2026polar};
  \item \emph{On-Policy Distillation of Language Models: Learning from
        Self-Generated Mistakes} \citep{agarwal2024onpolicy};
  \item \emph{AgentPRM} \citep{xi2025agentprm};
  \item \emph{Self-Distillation Enables Continual Learning}
        \citep{shenfeld2026selfdistillation};
  \item \emph{Let's Verify Step by Step} \citep{lightman2023verify};
  \item the \emph{GLM-5} technical report \citep{glm5team2026};
  \item the \emph{GLM-5.2} research article \citep{glm52team2026}; and
  \item the \emph{Composer 2.5} research article \citep{cursor2026composer25}.
\end{enumerate}
These source types are not treated as interchangeable evidence. Peer review, disclosure
depth, reproducibility, ablations, benchmark independence, and commercial incentives
will be evaluated explicitly. Nevertheless, technical reports are primary evidence for
important systems that are not yet described by conventional conference papers, so they
belong in the central corpus rather than in a footnote or an auxiliary reading list.

\subsection{Working position}
State the paper's eventual thesis here. A candidate structure is that increasingly
long-horizon agent training requires a deliberate combination of outcome signals,
localized process feedback, learned critics, and on-policy distribution matching rather
than a single universal post-training objective.
