\section{Transformer foundations: architecture, training, and inference}

The Transformer replaced recurrent sequence processing with attention layers that can
process all known sequence positions in parallel \citep{vaswani2017attention}. The
original paper studied an encoder--decoder architecture for translation. Modern causal
language models commonly retain only the decoder-style, autoregressive path. This section
uses that decoder-only form because it is the policy optimized by SFT, RL, and
distillation later in this paper.

\subsection{Representing tokens and positions}

For a batch of $B$ sequences of length $L$, tokenization produces integer IDs
$X\in\mathbb{N}^{B\times L}$. An embedding table maps each ID to a $D$-dimensional
vector. Self-attention alone has no direct representation of token distance, so position
information must also enter the computation. A basic implementation adds learned token
and position embeddings,
\begin{equation}
  H^{(0)}_{b,t}=E_{\mathrm{token}}[X_{b,t}]+E_{\mathrm{pos}}[t],
\end{equation}
giving $H^{(0)}\in\mathbb{R}^{B\times L\times D}$. The original Transformer used fixed
sinusoidal positions; many current LLMs instead modify queries and keys with rotary
position embeddings. The purpose is the same: two occurrences of the same token at
different positions must not enter attention as identical events.

\subsection{Causal multi-head self-attention}

One attention layer projects each token representation into queries, keys, and values:
\begin{equation}
  Q=HW_Q,\qquad K=HW_K,\qquad V=HW_V.
\end{equation}
With $h$ heads and head width $d_h=D/h$, these tensors are reshaped to
$[B,h,L,d_h]$. For one head, scaled dot-product attention is
\begin{equation}
  A=\operatorname{softmax}\!\left(\frac{QK^\top}{\sqrt{d_h}}+M\right),
  \qquad C=AV,
\end{equation}
where $M_{ij}=-\infty$ when key position $j$ is later than query position $i$.
The allowed attention pattern is lower triangular:
\begin{center}
\begin{tabular}{c|ccccc}
  & $k_0$ & $k_1$ & $k_2$ & $k_3$ & $k_4$ \\
  \hline
  $q_0$ & $\checkmark$ & & & & \\
  $q_1$ & $\checkmark$ & $\checkmark$ & & & \\
  $q_2$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & & \\
  $q_3$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & \\
  $q_4$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$ & $\checkmark$
\end{tabular}
\end{center}
Thus the representation at position $t$ can depend on $x_{\leq t}$ but not on future
tokens. Each head can learn a different content-dependent retrieval pattern. Their
contexts are concatenated and projected back to width $D$.

Attention is only one sublayer. A common pre-normalized decoder block is
\begin{align}
  H' &= H+\operatorname{Attention}(\operatorname{LayerNorm}(H)),\\
  H^{\mathrm{next}} &= H'+\operatorname{MLP}(\operatorname{LayerNorm}(H')).
\end{align}
Layer normalization stabilizes the input scale of each sublayer, residual connections
preserve an identity path through a deep network, and the token-wise MLP performs a
learned nonlinear transformation. The original Transformer used post-normalization;
the pre-normalized form above is easier to optimize at depth and matches the accompanying
implementation notebook.

After the final block, a normalized hidden vector at every position is projected to
$V_{\mathrm{vocab}}$ logits. Softmax converts those logits into the distribution of the
next token. The implementation path is therefore
\begin{center}
\texttt{IDs $\rightarrow$ embeddings+positions $\rightarrow$ repeated decoder blocks
$\rightarrow$ vocabulary logits}.
\end{center}

\subsection{Why training retains every token position}

Let $y=(y_1,\ldots,y_L)$ be one trajectory following a prompt $x$. An autoregressive
model does not emit one indivisible trajectory score. It factorizes the probability into
one decision per token:
\begin{equation}
  p_\theta(y\mid x)=\prod_{t=1}^{L}p_\theta(y_t\mid x,y_{<t}),
  \qquad
  \log p_\theta(y\mid x)=\sum_{t=1}^{L}\log p_\theta(y_t\mid x,y_{<t}).
\end{equation}
The batch dimension $B$ counts independent trajectories; the token dimension $L$ counts
the causally ordered decisions inside each trajectory. Replacing $[B,L,V]$ logits with
one $[B,V]$ output would retain only one next-token prediction per trajectory and discard
the remaining dense supervision.

For example, the sequence
\begin{center}
\texttt{[BOS, I, like, tea, EOS]}
\end{center}
defines four conditional examples: predict \texttt{I} after \texttt{BOS}, predict
\texttt{like} after \texttt{BOS I}, predict \texttt{tea} after \texttt{BOS I like},
and predict \texttt{EOS} after the complete response. Running these four prefixes
separately would repeat the early-token computation. Teacher forcing instead supplies
the complete correct sequence and uses the causal mask to compute every prefix-conditioned
prediction in one parallel forward pass.

The scale of this saving is visible at $L=4096$. Replaying every prefix, from length one
through 4096, would evaluate
\begin{equation}
  \sum_{t=1}^{4096}t=8{,}390{,}656
\end{equation}
token positions in projection and MLP layers. A teacher-forced pass evaluates 4096 token
positions once, a factor of 2048.5 fewer token-local evaluations. Dense self-attention
still performs quadratic pairwise interactions; parallel training avoids repeated prefix
forwards rather than eliminating the attention algorithm's $O(L^2D)$ arithmetic.

For selected target positions $\mathcal{S}$, next-token training minimizes
\begin{equation}
  \mathcal{L}(\theta)
  =-\frac{1}{|\mathcal{S}|}\sum_{t\in\mathcal{S}}
    \log p_\theta(y_{t+1}\mid y_{\leq t}).
\end{equation}
The factor $1/|\mathcal{S}|$ does \emph{not} save computation. It normalizes the gradient
scale so that loss magnitude is comparable across batches and sequence lengths. The
computational saving comes from the single causal forward pass. In pretraining,
$\mathcal{S}$ normally includes all non-padding next-token targets; SFT later restricts
it to assistant-owned targets.

A basic training step is:
\begin{enumerate}
  \item tokenize and batch sequences into \texttt{input\_ids} with shape $[B,L]$;
  \item run the decoder once to obtain logits $[B,L,V_{\mathrm{vocab}}]$;
  \item align logits at $t$ with target IDs at $t+1$;
  \item compute masked cross-entropy over selected targets;
  \item backpropagate through all decoder blocks and update the parameters.
\end{enumerate}
If $P$ is the number of model parameters and $N$ the number of processed tokens, a useful
first-order estimate is roughly $2PN$ operations for a dense forward pass and $6PN$ for
forward plus backward, before attention, recomputation, and systems overheads are treated
more precisely.

\subsection{Inference: prefill and autoregressive decoding}

Training and inference execute the same learned functions under different information
constraints. During training, all ground-truth tokens are known. During inference, token
$y_t$ does not exist until the model has selected $y_{t-1}$. Inference therefore has two
phases.

\paragraph{Prompt prefill.}
The system and user prompt are already known, so all prompt positions are processed in
parallel. Only the final next-token distribution is needed, but every layer also retains
the keys and values produced for the prompt.

\paragraph{Cached decoding.}
The model selects one new token, embeds it at the next position, and computes its query,
key, and value at each layer. The new query attends to all earlier keys and values. The
new key and value are appended to the cache, after which the model can generate the next
token.

Caching is valid because causality makes past representations immutable: appending a
future token cannot change a previous position that was forbidden to attend to that
future. Without a cache, step $g$ would rerun the complete prefix of length $P+g$. With a
cache, the prompt is processed once and each decoding step evaluates token-local
projections and MLPs only for the new token.

Consider a prompt of $P=4096$ tokens followed by $G=1024$ generated tokens. Without a
cache, the number of token positions repeatedly passed through projections and MLPs is
\begin{equation}
  \sum_{g=0}^{G-1}(P+g)
  =GP+\frac{G(G-1)}{2}
  =4{,}718{,}080.
\end{equation}
Prefill plus cached decoding processes $P+G=5120$ token positions through those
token-local operations, a factor of 921.5 fewer. If a fully dense attention matrix were
also recomputed for every uncached prefix, this example would calculate approximately
21.83 billion attention-score entries, versus approximately 21.50 million for one dense
prefill followed by one new-query row per decode step. These ratios explain avoided
recomputation, not an equal end-to-end latency improvement: kernels, memory traffic,
sampling, and other model operations also matter.

The cache has a cost. Per layer its approximate storage is
\begin{equation}
  2\,B\,H_{\mathrm{kv}}\,L\,d_h\,s_{\mathrm{dtype}}\ \text{bytes},
\end{equation}
where the leading 2 represents keys and values. At length 5120 with 32 KV heads, head
width 128, and two-byte values, this is 80 MiB per sequence per layer, or 2.5 GiB across
32 layers. Multi-query and grouped-query attention reduce $H_{\mathrm{kv}}$ and therefore
cache size and memory bandwidth \citep{shazeer2019fastdecoding}.

KV caching removes repeated K/V projection and MLP work for the history; it does not make
decoding constant-time. Each new query must still read a growing cache, so attention work
accumulates quadratically over a long generated sequence. Training normally does not use
decode-style caching because known tokens can be processed more efficiently in parallel
and backpropagation requires the layer activations.

\subsection{Training and inference as one system}

The distinction becomes especially important in post-training. SFT performs a parallel
training pass over demonstrations. RL first uses cached inference to sample trajectories,
then replays the now-known token IDs through a training engine to obtain differentiable
log probabilities. The rollout cache is not a substitute for this training forward pass:
it was commonly created without a gradient graph, possibly by an older policy and a
separate inference engine. Understanding a Transformer therefore requires both its
parallel training behavior and its stateful autoregressive inference behavior.
