\section{Background: learning signals after pretraining}

Pretraining gives a language model a broad ability to continue text. It does not by
itself specify which behavior should be preferred when the model follows instructions,
solves a verifiable problem, uses tools, or acts across a long interaction. Post-training
is the stage in which targeted data and feedback reshape those behaviors. The architecture
may remain unchanged; what changes is the experience presented to the model and the
learning signal extracted from it.

This paper studies post-training through one central question: what information does a
training signal provide about the behavior that should change? A demonstration supplies
desired token actions. A terminal reward judges a sampled response without identifying
which decision caused the result. A process evaluator attempts to judge intermediate
reasoning or progress. A teacher distribution supplies dense preferences over possible
next tokens. These signals are not interchangeable. They differ in what they reveal, the
model states on which they are available, the cost of obtaining them, and the errors they
can introduce.

\subsection{The trajectory is the unit that connects the methods}

A decoder-only language model generates one token at a time, but useful behavior is
usually judged over a larger trajectory. For a short question, the trajectory may be one
response. For an agent, it may contain many model calls, tool actions, environment
observations, context replacement, and delegated work. The training problem is therefore
not only whether a final answer was good. It is also how evidence about that outcome
should be assigned to the token decisions that produced it.

The difficulty grows with the horizon. A terminal reward can identify a successful
trajectory while saying little about which early step enabled success, which detour was
harmless, or where recovery became impossible. Tool responses provide essential context
but are normally not policy actions. Context compaction and parallel branches can also
break the assumption that one continuous token prefix was generated by one policy call.
Work on process supervision and agentic trace construction addresses different parts of
this problem \citep{lightman2023verify,xu2026polar}.

Long trajectories also turn training-system details into algorithmic concerns. Rollouts
are expensive and sequential, so systems overlap generation with optimization or reuse
stored responses. The learner may then update while workers still generate with an older
policy. Importance ratios, policy-version records, masking, and trace boundaries determine
whether the resulting update still corresponds to the actions that were actually sampled
\citep{glm5team2026}. For long-horizon post-training, data provenance is part of the
learning problem rather than merely an implementation detail.

\subsection{A spectrum of supervision}

Supervised fine-tuning provides the most direct target: imitate demonstrated response
tokens. It is effective when the desired behavior can be written down, but teacher forcing
mostly exposes the model to expert prefixes rather than to states created by its own
mistakes. SFT therefore teaches demonstrated behavior more directly than it teaches
exploration, outcome comparison, or recovery outside the demonstration distribution.

Outcome-reward reinforcement learning takes the complementary approach. The model samples
its own attempts, and a verifier or reward model judges the completed result. This places
training on the model's behavior distribution and can reward solutions that were never
demonstrated. The price is sparse credit assignment: one scalar judgment may be broadcast
across a long response. GRPO is an important instance because it estimates a reward
baseline by comparing sibling responses to the same prompt rather than training a
separate value model \citep{guo2025deepseekr1}.

Process reward models and critics try to add information between the initial prompt and
the terminal outcome. They can distinguish a locally incorrect step from a promising
partial trajectory, which becomes increasingly valuable as tasks lengthen. Their benefit
depends on whether the evaluator actually recognizes progress rather than merely producing
dense but unreliable scores. A learned critic also introduces its own training data,
inference cost, calibration problem, and opportunities for exploitation.

Distillation supplies another kind of dense signal. Instead of returning one score, a
teacher provides a distribution over next-token choices. On-policy distillation evaluates
that teacher on prefixes sampled by the student, combining dense token-level feedback with
the student's own state distribution \citep{lu2025onpolicydistillation}. This may teach
local corrections that terminal rewards cannot express, but its behavior is constrained by
the teacher's knowledge, the compatibility of teacher and student tokenizations, and the
contexts made available to the teacher.

The practical question is consequently not which method is universally best. It is which
signal is available for the target capability and what information is missing from that
signal. Demonstrations are appropriate when desired actions are known. Verifiable rewards
are attractive when outcomes are reliable but solution paths are diverse. Process
feedback becomes useful when intermediate progress matters. Distillation is attractive
when a stronger teacher can evaluate the student's encountered states. Production systems
may combine these methods because each repairs a different information deficit.

\subsection{From post-training stages to continual improvement}

The longer-term goal of this study is to understand post-training as a potentially
repeatable learning process rather than a single final stage. A system might synthesize
new tasks, attempt them, verify outcomes, diagnose failure modes, regenerate targeted
data, and construct the next curriculum. Repeating this loop resembles continual learning
and automated domain adaptation, but automation alone does not guarantee cumulative
improvement. The system must preserve earlier capabilities, avoid training on its own
evaluation artifacts, and distinguish genuine progress from reward exploitation.

Continual self-distillation offers one possible mechanism for transferring behavior across
successive model states \citep{shenfeld2026selfdistillation}. Outcome RL, critics, and
automated verification offer others. Whether their combination supports recursive
improvement is a research question for the literature and experiments, not an assumption
of this paper.

\subsection{Organizing questions and working position}

The paper uses five questions to organize the technical material and source review:
\begin{enumerate}
  \item What behavior can SFT, outcome-reward RL, process supervision, and distillation
        each teach or preserve?
  \item When does terminal reward cease to provide adequate credit for a long trajectory?
  \item What additional information must a process reward model or critic estimate, and
        how can that estimate fail?
  \item How does training on the model's own state distribution change generalization,
        correction, and recovery?
  \item Which data-generation, verification, and retention mechanisms are required for
        post-training to become a reliable continual-learning loop?
\end{enumerate}

The working position is that post-training methods should be compared by their source of
supervision, credit granularity, sampling distribution, evaluator reliability, and systems
constraints. Longer-horizon agents are unlikely to be served by one universal objective:
they require deliberate combinations of demonstrations, outcome signals, localized
feedback, distribution matching, and capability-retention mechanisms. The next section
begins with the decoder-only Transformer because every later method ultimately changes the
probabilities produced by that same autoregressive computation.
